Modeling site-specific amino-acid preferences deepens phylogenetic estimates of the divergence of viral proteins.

### 2018-03-26 manuscript/manuscript.tex
updated asymp. seq identity in manuscript

\ref{fig:decay} shows how different substitution models predict amino-acid sequence identity to decrease as a function of branch length using model parameters fit to a phylogeny of H1 influenza hemagglutinin (HA) genes.
The GY94 model predicts the same behavior for all sites, since it does not have any site-specific parameters, with an asymptotic sequence identity of 0.059. 
While this predicted sequence identity is higher than $\frac{1}{20} = 0.05$, due to redundant codon and nucleotide biases favoring certain amino acids, it is much lower than the pairwise identity of even the most diverged HAs in nature.
While it is of course possible that the identity of HAs in nature would become even lower given more time, it seems biochemically improbable that it would ever become as low as 0.059. 
The reason is that like many proteins HA has a highly conserved structure and function that imposes constraints that cause many sites to sample only a small subset of the 20 amino acids among all known HA homologs~\citep{nobusawa1991comparison}.

>>>>

\ref{fig:decay} shows how different substitution models predict amino-acid sequence identity to decrease as a function of branch length using model parameters fit to a phylogeny of H1 influenza hemagglutinin (HA) genes.
The GY94 model predicts the same behavior for all sites, since it does not have any site-specific parameters, with an asymptotic sequence identity of 0.062. 
While this predicted sequence identity is higher than $\frac{1}{20} = 0.05$, due to redundant codon and nucleotide biases favoring certain amino acids, it is much lower than the pairwise identity of even the most diverged HAs in nature.
While it is of course possible that the identity of HAs in nature would become even lower given more time, it seems biochemically improbable that it would ever become as low as 0.062. 
The reason is that like many proteins HA has a highly conserved structure and function that imposes constraints that cause many sites to sample only a small subset of the 20 amino acids among all known HA homologs~\citep{nobusawa1991comparison}.
